NAME: ZHIBO WANG
---

title: 'Project 2: Modeling and Evaluation'
subtitle: "CSE6242 - Data and Visual Analytics - Fall 2017\n\nDue: Sunday, April 22, 2018 at 11:59 PM UTC-12:00 on T-Square"
output:
  pdf_document: default
  html_document:
    code_folding: none
    theme: default
  html_notebook:
    code_folding: none
    theme: default
---

# Data

We will use the same dataset as Project 1: [`movies_merged`](https://s3.amazonaws.com/content.udacity-data.com/courses/gt-cs6242/project/movies_merged).

# Objective

Your goal in this project is to build a linear regression model that can predict the `Gross` revenue earned by a movie based on other variables. You may use R packages to fit and evaluate a regression model (no need to implement regression yourself). Please stick to linear regression, however.

# Instructions

You should be familiar with using an [RMarkdown](http://rmarkdown.rstudio.com) Notebook by now. Remember that you have to open it in RStudio, and you can run code chunks by pressing *Cmd+Shift+Enter*.

Please complete the tasks below and submit this R Markdown file (as **pr2.Rmd**) containing all completed code chunks and written responses, and a PDF export of it (as **pr2.pdf**) which should include the outputs and plots as well.

_Note that **Setup** and **Data Preprocessing** steps do not carry any points, however, they need to be completed as instructed in order to get meaningful results._

# Setup

Same as Project 1, load the dataset into memory:

```{r, echo=FALSE, warning=FALSE}
load('movies_merged')
```

This creates an object of the same name (`movies_merged`). For convenience, you can copy it to `df` and start using it:

```{r}
df = movies_merged
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
colnames(df)
```

## Load R packages

Load any R packages that you will need to use. You can come back to this chunk, edit it and re-run to load any additional packages later.

```{r}
#install.packages("tidyr")

library(ggplot2)
library(stringr)
library(tidyr)
```

If you are using any non-standard packages (ones that have not been discussed in class or explicitly allowed for this project), please mention them below. Include any special instructions if they cannot be installed using the regular `install.packages('<pkg name>')` command.

**Non-standard packages used**: None

# Data Preprocessing

Before we start building models, we should clean up the dataset and perform any preprocessing steps that may be necessary. Some of these steps can be copied in from your Project 1 solution. It may be helpful to print the dimensions of the resulting dataframe at each step.

## 1. Remove non-movie rows

```{r}
# TODO: Remove all rows from df that do not correspond to movies
df2 <- df[df$Type == "movie",]
```

## 2. Drop rows with missing `Gross` value

Since our goal is to model `Gross` revenue against other variables, rows that have missing `Gross` values are not useful to us.

```{r}
# TODO: Remove rows with missing Gross value
df2<-df2[!is.na(df2$Gross),]
```

## 3. Exclude movies released prior to 2000

Inflation and other global financial factors may affect the revenue earned by movies during certain periods of time. Taking that into account is out of scope for this project, so let's exclude all movies that were released prior to the year 2000 (you may use `Released`, `Date` or `Year` for this purpose).

```{r}
# TODO: Exclude movies released prior to 2000


rmlist<-NULL
totaln<- length(df2$Year)
for (n in 1:totaln){
   if ((as.numeric(df2$Year[n]))<2000)
   rmlist<-c(rmlist,n)
}
if (!is.null(rmlist))
  df2<-df2[-c(rmlist),]

```

## 4. Eliminate mismatched rows

_Note: You may compare the `Released` column (string representation of release date) with either `Year` or `Date` (numeric representation of the year) to find mismatches. The goal is to avoid removing more than 10% of the rows._

```{r}
# TODO: Remove mismatched rows
rmlist<-NULL
for (i in 1:length(df2$Title)){
  #step1. parsed Year, Date and Relaesed columns and get numeric value of years
  year <- as.numeric(df2$Year[i])
  date <- as.numeric(df2$Date[i])
  release <- as.numeric(format(df2$Released[i], format="%Y"))
 #step2. for each movie, need to check 3 pairs of year values for consistency:
  #(Date,Year)
  #(Date,Released)
  #(Year, Released)
  #if there is any "na"" value in a particular pair, then assume that pair is consistent;if the difference in the pair is greater than +/- one year,  then assume that pair is inconsistent; if movie has at least one inconsistent pairs, then that movie shall be removed

  if (!(is.na(year) ||  is.na(release))){
    if (( (year+1) < release ) || ((year-1) > release) )
      rmlist<-c(rmlist,i)}
  else if (!(is.na(date) ||  is.na(release))){
    if (( (date+1) < release ) || ((date-1) > release) )
      rmlist<-c(rmlist,i)}
  else if (!(is.na(date) ||  is.na(year))){
    if (( (date+1) < year ) || ((date-1) > year) )
      rmlist<-c(rmlist,i)}

}
#step3. remove rows in dataframe according to "rmlist":
rmlist <- unique(rmlist )
if (!is.null(rmlist))
  df2<-df2[-c(rmlist),]
```

## 5. Drop `Domestic_Gross` column

`Domestic_Gross` is basically the amount of revenue a movie earned within the US. Understandably, it is very highly correlated with `Gross` and is in fact equal to it for movies that were not released globally. Hence, it should be removed for modeling purposes.

```{r}
# TODO: Exclude the `Domestic_Gross` column
df2<-df2[,-37]
```

## 6. Process `Runtime` column

```{r}
# TODO: Replace df$Runtime with a numeric column containing the runtime in minutes

#first clear n/a values, if runtime is n/a , remove row:
nalist<-NULL
totaln<- length(df2$Runtime)
for (n in 1:totaln){
   if (df2$Runtime[n]=="N/A")
   nalist<-c(nalist,n)
}
if (!is.null(nalist))
  df2<-df2[-c(nalist),]
# now df2 does not contain n/a runtime

#next change runtime from string to numeric:
totaln<- length(df2$Runtime)
for (n in 1:totaln){
  
  nruntime <- unlist(strsplit(df2$Runtime[n], " "))
  
  if (is.element("h",nruntime) && is.element("min",nruntime)){
    nhour <- as.numeric(nruntime[1])
    nmin <- as.numeric(nruntime[3])
    totalmin <- nhour*60+nmin
  }  else if (is.element("h",nruntime) && (!is.element("min",nruntime))) {
    nhour <- as.numeric(nruntime[1])
    totalmin <- nhour*60
  } else {
    nmin <- as.numeric(nruntime[1])
    totalmin <- nmin
  }
  
  df2$Runtime[n]<- totalmin
  
}
df2$Runtime <- as.numeric(as.character(df2$Runtime))

```

Perform any additional preprocessing steps that you find necessary, such as dealing with missing values or highly correlated columns (feel free to add more code chunks, markdown blocks and plots here as necessary).

```{r}
# TODO(optional): Additional preprocessing
#remove "na" imdb and tomato reviews and ratings

df2<-df2[!is.na(df2$imdbRating),]
df2<-df2[!is.na(df2$imdbVotes),]
df2<-df2[!is.na(df2$tomatoUserReviews),]
df2<-df2[!is.na(df2$tomatoUserRating),]
#below removes any movie with zero gross
rmlist<-NULL
for (i in 1:length(df2[,1]) ) {
  if (df2[i,37]==0)
  rmlist<- c(rmlist, i)
}
df2<-df2[-rmlist, ] 

```

_**Note**: Do NOT convert categorical variables (like `Genre`) into binary columns yet. You will do that later as part of a model improvement task._

## Final preprocessed dataset

Report the dimensions of the preprocessed dataset you will be using for modeling and evaluation, and print all the final column names. (Again, `Domestic_Gross` should not be in this list!)

```{r}
# TODO: Print the dimensions of the final preprocessed dataset and column names
dim(df2)
names(df2)


```

# Evaluation Strategy

In each of the tasks described in the next section, you will build a regression model. In order to compare their performance, you will compute the training and test Root Mean Squared Error (RMSE) at different training set sizes.

First, randomly sample 10-20% of the preprocessed dataset and keep that aside as the **test set**. Do not use these rows for training! The remainder of the preprocessed dataset is your **training data**.

Now use the following evaluation procedure for each model:

- Choose a suitable sequence of training set sizes, e.g. 10%, 20%, 30%, ..., 100% (10-20 different sizes should suffice). For each size, sample that many inputs from the training data, train your model, and compute the resulting training and test RMSE.
- Repeat your training and evaluation at least 10 times at each training set size, and average the RMSE results for stability.
- Generate a graph of the averaged train and test RMSE values as a function of the train set size (%), with optional error bars.

You can define a helper function that applies this procedure to a given set of features and reuse it.

# Tasks

Each of the following tasks is worth 20 points, for a total of 100 points for this project. Remember to build each model as specified, evaluate it using the strategy outlined above, and plot the training and test errors by training set size (%).

## 1. Numeric variables

Use Linear Regression to predict `Gross` based on available _numeric_ variables. You can choose to include all or a subset of them.

```{r}
# TODO: Build & evaluate model 1 (numeric variables only)
################################################################
#first randomly select train dataset and test dataset 
selectrows<- function(data,percent){
  all<- seq.int(from = 1, to = length(data[,1])) 
  n<-as.integer(length(data[,1])*percent)
  select<-  sample(all, size = n)
  return(select)
}
set.seed(47)
trainrows<- selectrows(df2,0.9) # trainset uses 90% data, test set uses 10% data
trainset<- df2[trainrows,]
testset<- df2[-trainrows,]  

#################################################################################
#create a function to do training on the data
trainmodel1<- function(data){
    #model<- lm(Gross~Budget, data)
    #model<- lm(Gross~Budget+imdbRating, data)
    #model<- lm(Gross~Budget+imdbVotes, data)
    #model<- lm(Gross~Budget+imdbRating+imdbVotes+imdbRating*imdbVotes, data)
   model<- lm(Gross~Budget+Year+Runtime+imdbRating+imdbVotes+tomatoUserRating+tomatoUserReviews, data)
  return(model)
}
# models with j=20 various % of the data, repeat i=10 times for each %
trealgross<- testset[37]
trainrmse<- matrix( 1:40,  nrow=20, ncol=2) 
testrmse<- matrix( 1:40,  nrow=20, ncol=2) 
rmse<- 0
for (j in 1:20){ # use 20 different percentages
  lista<-NULL
  listb<-NULL
  for (i in 1:20){ # repeat 20 times
  select<-selectrows(trainset,j/20)
  data<- trainset[select,]
  M1<- trainmodel1(data)
  rmse<-sqrt(mean(residuals(M1)^2))
  lista<-c(lista,rmse)
  tpgross<- as.data.frame( predict(M1,testset))
  trmse<-0
  rss<- 0
  for (a in 1:length(testset[,1])){
  e<- (tpgross[a,1]-trealgross[a,1])^2
  rss<- rss+e
  }
  trmse<- sqrt(rss/length(testset[,1]))
  listb<-c(listb,trmse)
  }
  trainrmse[j,2]<-mean(lista) # this records all train rmse
  trainrmse[j,1]<- j*5 # this records the percentage of data used
  testrmse[j,2]<-mean(listb) # this records all train rmse
  testrmse[j,1]<- j*5 # this records the percentage of data used
}
testresult1<- as.data.frame(testrmse)
colnames(testresult1)<- c("percentage","RMSE")
trainresult1<- as.data.frame(trainrmse)
colnames(trainresult1)<- c("percentage","RMSE")
plotgraph<-function(testresult1,trainresult1){
ggplot(testresult1,aes(y=RMSE, x = percentage)) +  geom_point(data=testresult1,aes(y=RMSE, x = percentage, colour='test')) +  geom_line(data=testresult1,aes(y=RMSE, x = percentage, colour='test')) + geom_point(data=trainresult1,aes(y=RMSE, x = percentage, colour='train')) +  geom_line(data=trainresult1,aes(y=RMSE, x = percentage, colour='train'))
}

plotgraph(testresult1,trainresult1)

```

**Q**: List the numeric variables you used.

**A**: variables used: Budget+Year+Runtime+imdbRating+imdbVotes+tomatoUserRating+tomatoUserReviews


**Q**: What is the best mean test RMSE value you observed, and at what training set size?

**A**: best mean test RMSE=86002539 , trained with 90% training set


## 2. Feature transformations

Try to improve the prediction quality from **Task 1** as much as possible by adding feature transformations of the numeric variables. Explore both numeric transformations such as power transforms and non-numeric transformations of the numeric variables like binning (e.g. `is_budget_greater_than_3M`).

```{r}
# TODO: Build & evaluate model 2 (transformed numeric variables only)
#create a function to do training on the data
testset2<-testset
trainset2<-trainset

#bin budget into 4 category: 0 means lowbudget(<3M), 1 means midbudget(>3M but < 10M), 2 means highbudget (>10M but <100M) and 3 means superbudget(>100M) 
for (x in 1:length(testset2[,36])) {
  if (testset2[x,36]<=3000000)
    testset2[x,39]<- 0
  if (testset2[x,36]>3000000 && testset2[x,36]<=10000000)
    testset2[x,39]<- 1
  if (testset2[x,36]>10000000 && testset2[x,36]<=100000000)
    testset2[x,39]<- 2
  if (testset2[x,36]>100000000)
    testset2[x,39]<- 3
}

for (x in 1:length(trainset2[,36])) {
    if (trainset2[x,36]<=3000000)
    trainset2[x,39]<- 0
  if (trainset2[x,36]>3000000 && trainset2[x,36]<=10000000)
    trainset2[x,39]<- 1
  if (trainset2[x,36]>10000000 && trainset2[x,36]<=100000000)
    trainset2[x,39]<- 2
  if (trainset2[x,36]>100000000)
    trainset2[x,39]<- 3
}
colnames(testset2)[39]<- "budgetbin"
colnames(trainset2)[39]<- "budgetbin"

trainmodel2<- function(data){
    #model<- lm(Gross~Budget, data)
    #model<- lm(Gross~Budget+imdbRating, data)
    #model<- lm(Gross~Budget+imdbVotes, data)
    #model<- lm(Gross~Budget+imdbRating+imdbVotes+imdbRating*imdbVotes, data)
   #model<- lm(Gross~Budget+log(Budget)+Year+Runtime+log(Runtime)+imdbRating+imdbVotes+tomatoUserRating+tomatoUserReviews, data)
   model<- lm(Gross~Budget+I(Budget^2)+I(Budget^3)+Year+Runtime+I(log(Runtime))+imdbRating*imdbVotes+tomatoUserRating*tomatoUserReviews+budgetbin, data)
  return(model)
}
# models with j=20 various % of the data, repeat i=10 times for each %
trealgross<- testset[37]
trainrmse<- matrix( 1:40,  nrow=20, ncol=2) 
testrmse<- matrix( 1:40,  nrow=20, ncol=2) 
rmse<- 0
for (j in 1:20){ # use 19 different percentages
  lista<-NULL
  listb<-NULL
  for (i in 1:20){ # repeat 20 times
  select<-selectrows(trainset2,j/20)
  data<- trainset2[select,]
  M2<- trainmodel2(data)
  rmse<-sqrt(mean(residuals(M2)^2))
  lista<-c(lista,rmse)
  tpgross<- as.data.frame( predict(M2,testset2))
  trmse<-0
  rss<- 0
  for (a in 1:length(testset2[,1])){
  e<- (tpgross[a,1]-trealgross[a,1])^2
  rss<- rss+e
  }
  trmse<- sqrt(rss/length(testset2[,1]))
  listb<-c(listb,trmse)
  }
  trainrmse[j,2]<-mean(lista) # this records all train rmse
  trainrmse[j,1]<- j*5 # this records the percentage of data used
  testrmse[j,2]<-mean(listb) # this records all train rmse
  testrmse[j,1]<- j*5 # this records the percentage of data used
}
testresult2<- as.data.frame(testrmse)
colnames(testresult2)<- c("percentage","RMSE")
trainresult2<- as.data.frame(trainrmse)
colnames(trainresult2)<- c("percentage","RMSE")

plotgraph(testresult2,trainresult2)

###########################################################################

```

**Q**: Explain which transformations you used and why you chose them.

**A**: 
1. Budget  vs  Gross plot is not very linear and shows slight second and third order trend, therefore modeled using Budget+I(Budget^2)+I(Budget^3)

2. Runtime vs Gross shows slightl linear relationship in log scale, therefore used Runtime+I(log(Runtime))

3. rating and number of reviews are not very good predictors when considered alone, some highly rated movies are not popular and only have a few reviews; on the other hand, movies with many reviews may have very low ratings.  Using rating*number of reviews will compensate for this, therefore imdbRating*imdbVotes+tomatoUserRating*tomatoUserReviews are used


**Q**: How did the RMSE change compared to Task 1?

**A**: Overall the RMSE is slightly reduced. The lowest RMSE reduced to 85509369 (with 100% train data)



## 3. Non-numeric variables

Write code that converts genre, actors, directors, and other categorical variables to columns that can be used for regression (e.g. binary columns as you did in Project 1). Also process variables such as awards into more useful columns (again, like you did in Project 1). Now use these converted columns only to build your next model.

```{r, echo=FALSE, warning=FALSE}
# TODO: Build & evaluate model 3 (converted non-numeric variables only)
#backup and save separate test and train data as below:
testset3<-testset2
trainset3<-trainset2
df3<-df2
#################################################################################
#split award into wins and nominations
processaward<- function(df){
wins <- as.numeric(str_extract(str_extract(df$Awards, '\\d+ wins?'), '\\d+'))
wins[is.na(wins)] <- 0
nominations <- as.numeric(str_extract(str_extract(df$Awards, '\\d+ nominations?'), '\\d+'))
nominations[is.na(nominations)] <- 0
won <- as.numeric(str_extract(str_extract(df$Awards, 'Won \\d+'), '\\d+'))
won[is.na(won)] <- 0
nominated <- as.numeric(str_extract(str_extract(df$Awards, 'Nominated for \\d+'), '\\d+'))
nominated[is.na(nominated)] <- 0
na <- df$Awards == 'N/A'
df$Wins <- wins + won
df$Wins[na] <- 0 #na become 0
df$Nominations <- nominations + nominated
df$Nominations[na] <- 0 #na become 0
return(df)
}

testset3<- processaward(testset3)
trainset3<- processaward(trainset3)
#################################################################################
#split genre
process<- function(dfa,dfb){
glist <- NULL
totaln <- length(dfa$Genre)
for (n in 1:totaln){
    gen <- unlist(strsplit(dfa$Genre[n], ", "))
    glist<- (c(glist,gen))
}

glist<-names(sort(summary(as.factor(glist)), decreasing=T)[1:11])
glist<-glist[-1]


# for every genre in glist, create a new colum in dataframe:
for (g in glist){
  colname <- g
  genrevalue<- NULL
  for (i in 1:length(dfb$Genre)){
    if (grepl(g, dfb$Genre[i], fixed=TRUE))
      genrevalue <- c(genrevalue,1)
    else genrevalue <- c(genrevalue,0)
  }
  dfb[colname] <- genrevalue
}
return(dfb)
}

trainset3<- process(df3,trainset3)
testset3<- process(df3,testset3)

#################################################################################
#split Director
process<- function(dfa,dfb){
glist <- NULL
totaln <- length(dfa$Director)
for (n in 1:totaln){
    gen <- unlist(strsplit(dfa$Director[n], ", "))
    glist<- (c(glist,gen))
}

glist<-names(sort(summary(as.factor(glist)), decreasing=T)[1:11])
glist<-glist[-1]



# for every genre in glist, create a new colum in dataframe:
for (g in glist){
  colname <- g
  genrevalue<- NULL
  for (i in 1:length(dfb$Director)){
    if (grepl(g, dfb$Director[i], fixed=TRUE))
      genrevalue <- c(genrevalue,1)
    else genrevalue <- c(genrevalue,0)
  }
  dfb[colname] <- genrevalue
}
return(dfb)
}

trainset3<- process(df3,trainset3)
testset3<- process(df3,testset3)

#################################################################################
#split Actors
process<- function(dfa,dfb){
glist <- NULL
totaln <- length(dfa$Actors)
for (n in 1:totaln){
    gen <- unlist(strsplit(dfa$Actors[n], ", "))
    glist<- (c(glist,gen))
}

glist<-names(sort(summary(as.factor(glist)), decreasing=T)[1:11])
glist<-glist[-1]


# for every genre in glist, create a new colum in dataframe:
for (g in glist){
  colname <- g
  genrevalue<- NULL
  for (i in 1:length(dfb$Actors)){
    if (grepl(g, dfb$Actors[i], fixed=TRUE))
      genrevalue <- c(genrevalue,1)
    else genrevalue <- c(genrevalue,0)
  }
  dfb[colname] <- genrevalue
}
return(dfb)
}

trainset3<- process(df3,trainset3)
testset3<- process(df3,testset3)


#################################################################################
names(testset3)<-gsub("\\s","",names(testset3))
names(testset3)<-gsub("-","",names(testset3))
names(trainset3)<-gsub("\\s","",names(trainset3))
names(trainset3)<-gsub("-","",names(trainset3))

#now data is ready for lm()
#create a function to do training on the data
trainmodel3<- function(data){
 model<- lm((paste("Gross", "~", paste(colnames(data)[c(40:71)], collapse = "+"), sep = "" )), data=data )
 #model using all columns, include wins, nominations, top 10 genre, top 10 director and top 10 actor
 return(model)
}
# models with j=20 various % of the data, repeat i=10 times for each %
trealgross<- testset3[37]
trainrmse<- matrix( 1:40,  nrow=20, ncol=2) 
testrmse<- matrix( 1:40,  nrow=20, ncol=2) 
rmse<- 0
for (j in 1:20){ # use 20 different percentages
  lista<-NULL
  listb<-NULL
  for (i in 1:20){ # repeat 20 times
  select<-selectrows(trainset3,j/20)
  data<- trainset3[select,]
  M3<- trainmodel3(data)
  rmse<-sqrt(mean(residuals(M3)^2))
  lista<-c(lista,rmse)
  tpgross<- as.data.frame( predict(M3,testset3))
  trmse<-0
  rss<- 0
  for (a in 1:length(testset3[,1])){
  e<- (tpgross[a,1]-trealgross[a,1])^2
  rss<- rss+e
  }
  trmse<- sqrt(rss/length(testset3[,1]))
  listb<-c(listb,trmse)
  }
  trainrmse[j,2]<-mean(lista) # this records all train rmse
  trainrmse[j,1]<- j*5 # this records the percentage of data used
  testrmse[j,2]<-mean(listb) # this records all train rmse
  testrmse[j,1]<- j*5 # this records the percentage of data used
}
testresult3<- as.data.frame(testrmse)
colnames(testresult3)<- c("percentage","RMSE")
trainresult3<- as.data.frame(trainrmse)
colnames(trainresult3)<- c("percentage","RMSE")

ggplot(testresult3,aes(y=RMSE, x = percentage)) +  geom_point(data=testresult3,aes(y=RMSE, x = percentage, colour='test')) +  geom_line(data=testresult3,aes(y=RMSE, x = percentage, colour='test')) + geom_point(data=trainresult3,aes(y=RMSE, x = percentage, colour='train')) +  geom_line(data=trainresult3,aes(y=RMSE, x = percentage, colour='train'))





```

**Q**: Explain which categorical variables you used, and how you encoded them into features.

**A**: 
1.encoded award into wins and nominations with binary values 0/1
2. encoded genre into 10 new columns corresponding to top 10 genre names, each column has with binary values 0/1. Only picked top 10 genres because I try to reduce total number of variables, so as to reduce overfitting problem
3. similar to genre, encoded actors to top 10 actors columns, encoded with binary 0/1 
4. similar to genre, encoded directors to top 10 directors, encoded with binary 0/1 


**Q**: What is the best mean test RMSE value you observed, and at what training set size? How does this compare with Task 2?

**A**: the best best RMSE is 145903729 at 100% train data, as train data size increase, the RMSE keeps reducing.  The TASK3 best RMSE is worse than TASK2.



## 4. Numeric and categorical variables

Try to improve the prediction quality as much as possible by using both numeric and non-numeric variables from **Tasks 2 & 3**.

```{r}
# TODO: Build & evaluate model 4 (numeric & converted non-numeric variables)


trainmodel4<- function(data){
  model<- lm((paste("Gross~Budget+Year+Runtime+I(Runtime^2)+I(Runtime^3)+imdbRating*imdbVotes+tomatoUserRating*tomatoUserReviews+budgetbin+", paste(colnames(data)[c(40:71)], collapse = "+"), sep = "" )), data=data )
 #model using all columns, include wins, nominations, top 10 genre, top 10 director and top 10 actor
 return(model)
}


# models with j=20 various % of the data, repeat i=10 times for each %
trealgross<- testset3[37]
trainrmse<- matrix( 1:40,  nrow=20, ncol=2) 
testrmse<- matrix( 1:40,  nrow=20, ncol=2) 
rmse<- 0
for (j in 1:20){ # use 20 different percentages
  lista<-NULL
  listb<-NULL
  for (i in 1:20){ # repeat 20 times
  select<-selectrows(trainset3,j/20)
  data<- trainset3[select,]
  M3<- trainmodel4(data)
  rmse<-sqrt(mean(residuals(M3)^2))
  lista<-c(lista,rmse)
  tpgross<- as.data.frame( predict(M3,testset3))
  trmse<-0
  rss<- 0
  for (a in 1:length(testset3[,1])){
  e<- (tpgross[a,1]-trealgross[a,1])^2
  rss<- rss+e
  }
  trmse<- sqrt(rss/length(testset3[,1]))
  listb<-c(listb,trmse)
  }
  trainrmse[j,2]<-mean(lista) # this records all train rmse
  trainrmse[j,1]<- j*5 # this records the percentage of data used
  testrmse[j,2]<-mean(listb) # this records all train rmse
  testrmse[j,1]<- j*5 # this records the percentage of data used
}
testresult3<- as.data.frame(testrmse)
colnames(testresult3)<- c("percentage","RMSE")
trainresult3<- as.data.frame(trainrmse)
colnames(trainresult3)<- c("percentage","RMSE")

ggplot(testresult3,aes(y=RMSE, x = percentage)) +  geom_point(data=testresult3,aes(y=RMSE, x = percentage, colour='test')) +  geom_line(data=testresult3,aes(y=RMSE, x = percentage, colour='test')) + geom_point(data=trainresult3,aes(y=RMSE, x = percentage, colour='train')) +  geom_line(data=trainresult3,aes(y=RMSE, x = percentage, colour='train'))



```

**Q**: Compare the observed RMSE with Tasks 2 & 3.

**A**: task 4 lowest RMSE is 83663584, improved compared to 2 nd 3


## 5. Additional features

Now try creating additional features such as interactions (e.g. `is_genre_comedy` x `is_budget_greater_than_3M`) or deeper analysis of complex variables (e.g. text analysis of full-text columns like `Plot`).

```{r}
# TODO: Build & evaluate model 5 (numeric, non-numeric and additional features)



trainmodel4<- function(data){
  model<- lm((paste("Gross~Budget*imdbRating+ Budget+Year+Runtime+I(Runtime^2)+I(Runtime^3)+imdbRating*imdbVotes+tomatoUserRating*tomatoUserReviews+budgetbin+", paste(colnames(data)[c(40:71)], collapse = "+"), sep = "" )), data=data )
 #model using all columns, include wins, nominations, top 10 genre, top 10 director and top 10 actor
 return(model)
}


# models with j=20 various % of the data, repeat i=10 times for each %
trealgross<- testset3[37]
trainrmse<- matrix( 1:40,  nrow=20, ncol=2) 
testrmse<- matrix( 1:40,  nrow=20, ncol=2) 
rmse<- 0
for (j in 1:20){ # use 20 different percentages
  lista<-NULL
  listb<-NULL
  for (i in 1:20){ # repeat 20 times
  select<-selectrows(trainset3,j/20)
  data<- trainset3[select,]
  M3<- trainmodel4(data)
  rmse<-sqrt(mean(residuals(M3)^2))
  lista<-c(lista,rmse)
  tpgross<- as.data.frame( predict(M3,testset3))
  trmse<-0
  rss<- 0
  for (a in 1:length(testset3[,1])){
  e<- (tpgross[a,1]-trealgross[a,1])^2
  rss<- rss+e
  }
  trmse<- sqrt(rss/length(testset3[,1]))
  listb<-c(listb,trmse)
  }
  trainrmse[j,2]<-mean(lista) # this records all train rmse
  trainrmse[j,1]<- j*5 # this records the percentage of data used
  testrmse[j,2]<-mean(listb) # this records all train rmse
  testrmse[j,1]<- j*5 # this records the percentage of data used
}
testresult3<- as.data.frame(testrmse)
colnames(testresult3)<- c("percentage","RMSE")
trainresult3<- as.data.frame(trainrmse)
colnames(trainresult3)<- c("percentage","RMSE")

ggplot(testresult3,aes(y=RMSE, x = percentage)) +  geom_point(data=testresult3,aes(y=RMSE, x = percentage, colour='test')) +  geom_line(data=testresult3,aes(y=RMSE, x = percentage, colour='test')) + geom_point(data=trainresult3,aes(y=RMSE, x = percentage, colour='train')) +  geom_line(data=trainresult3,aes(y=RMSE, x = percentage, colour='train'))


```

